### 2015-03-30 paper/conclustion.tex
fixing some issues

\section{Conclusion}

We propose a scalable solution for DEBS grand challenge problem 2015. Our queries execute with O(log(n)) time complexity and hence scales with window size. We propose a scalable evaluation technique to execute queries in a distributed setup. Since \textit{EventEmitter} partitions the events using the  hash code value of the key (\textit{Route} or \textit{Cell}), our solution can be used to process events with multiple \textit{EventEmitters} (multiple event streams) as well. Different levels of \textit{ResultAggregators} can be used to scales up aggregation process if required. We propose an algorithm to find the runtime median with O(log(n)) time complexity. This algorithm specially suites for finding median with repeating values. The time windows used in our solution has an O(1) time complexity but O(n) space complexity. This creates opportunity to research further on efficient memory management techniques for large window sizes.